---
title: ""
date: ""
output:
  pdf_document:
    latex_engine: pdflatex
    number_sections: true
    toc: false # We will place it manually
header-includes:
  - \usepackage{fancyhdr}
  - \usepackage{xcolor}
  - \usepackage{titlesec}
  - \usepackage{wallpaper}
  - \usepackage{graphicx}
  - \usepackage{tikz}
  - \usepackage{geometry}
  - \usepackage{hyperref}
  - \usepackage{fontawesome5}
  - \definecolor{headerblue}{RGB}{0, 51, 102}
  - \definecolor{subblue}{RGB}{51, 102, 153}
  - \definecolor{sidebarcolor}{RGB}{240, 244, 248}
  - \pagestyle{fancy}
  - \fancyhf{}
  - \renewcommand{\headrulewidth}{0.4pt}
  - \titleformat{\section}{\color{headerblue}\normalfont\Large\bfseries}{\thesection}{1em}{}
  - \titleformat{\subsection}{\color{subblue}\normalfont\large\bfseries}{\thesubsection}{1em}{}
params:
  exclusion_log: NULL
  version: "1.3.0"
  img_path: "/srv/shiny-server/www/ppin.png"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
log <- params$exclusion_log
```

\begin{titlepage}
    \newgeometry{left=2cm,right=2cm,top=2cm,bottom=2cm}
    \begin{tikzpicture}[remember picture,overlay]
        % Top decorative bar
        \fill[headerblue] (current page.north west) rectangle ([yshift=-1cm]current page.north east);
        
        % Main Title
        \node[anchor=north west, text=headerblue] at ([xshift=2cm, yshift=-3cm]current page.north west) {
            \parbox{17cm}{
                {\fontsize{34}{40} \selectfont \bfseries olinkWrapper} \\
                \vspace{0.3cm}
                {\large \textcolor{headerblue!50}{\rule{17cm}{2pt}}} \\
                \vspace{0.3cm}
                {\Large \bfseries \color{subblue} Preprocessing Exclusion Report}
            }
        };

        % Center Figure (ppin.png)
        \node[anchor=center] at (current page.center) {
            \includegraphics[width=17cm]{`r params$img_path`}
        };

        % Bottom Metadata Block
        \node[anchor=south west] at ([xshift=2cm, yshift=5cm]current page.south west) {
            \parbox{17cm}{
                \begin{minipage}[t]{0.33\textwidth}
                    \textbf{\textcolor{headerblue}{Report Date}} \\
                    `r format(Sys.Date(), "%B %d, %Y")`
                \end{minipage}
                \begin{minipage}[t]{0.33\textwidth}
                    \textbf{\textcolor{headerblue}{Software Version}} \\
                    OlinkWrapper v`r params$version`
                \end{minipage}
                \begin{minipage}[t]{0.33\textwidth}
                    \textbf{\textcolor{headerblue}{App Developer}} \\
                    Jyotirmoy Das
                \end{minipage}
            }            
        };

        % Disclaimer
        \node[anchor=south west] at ([xshift=2cm, yshift=3.5cm]current page.south west) {
            \parbox{17cm}{
                \textbf{Disclaimer:} Please remember \textbf{olinkWrapper} is based on the \textcolor{headerblue}{\textbf{\href{https://github.com/Olink-Proteomics/OlinkRPackage}{OlinkAnalyze}}} package from the Olink Developers. Please cite/use the original package.
            }
        };
        

        % Bottom Citation
        \node[anchor=south west] at ([xshift=2cm, yshift=2.5cm]current page.south west) {
          \parbox{17cm}{
            \begin{minipage}[t]{0.5\textwidth}
              \textcolor{headerblue}{\faGlobe} {\href{https://olink-wrapper.serve.scilifelab.se/}{olink-wrapper.serve.scilifelab.se}}
            \end{minipage}
             \begin{minipage}[t]{0.5\textwidth}
              \textcolor{headerblue}{\faGithub} {\href{https://github.com/JD2112/olinkWrapper}{github.com/JD2112/olinkWrapper}}
            \end{minipage}
          }
        };

        \node[anchor=south west] at ([xshift=2cm, yshift=1.5cm]current page.south west) {
          \parbox{17cm}{            
              \small \textcolor{black}{Citation: \href{https://doi.org/10.5281/zenodo.15098644}{Zenodo: 15098644}}
          }
        };

        % Bottom decorative bar
        \fill[headerblue] (current page.south west) rectangle ([yshift=1cm]current page.south east);
    \end{tikzpicture}
\end{titlepage}

\restoregeometry
\newpage
\ClearWallPaper
\pagestyle{fancy}
\fancyhf{}

\fancyhead[C]{
  \begin{tikzpicture}[remember picture,overlay]
    \fill[headerblue] (current page.north west) rectangle ([yshift=-1.2cm]current page.north east);
    \node[anchor=north east, text=white] at ([xshift=-1cm, yshift=-0.4cm]current page.north east) {OlinkWrapper v`r params$version` };
  \end{tikzpicture}
}

\fancyfoot[C]{
  \begin{tikzpicture}[remember picture,overlay]
    \fill[headerblue] (current page.south west) rectangle ([yshift=1.2cm]current page.south east);
    \node[anchor=south west, text=white] at ([xshift=1cm, yshift=0.4cm]current page.south west) {\small \thepage};
    \node[anchor=south east, text=white] at ([xshift=-1cm, yshift=0.4cm]current page.south east) {\small Citation: \href{https://doi.org/10.5281/zenodo.15098644}{Zenodo: 15098644}};
  \end{tikzpicture}
}

\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}
\vspace*{1cm}

\tableofcontents
\newpage

# Introduction

This report summarises all exclusion steps performed during the preprocessing of Olink NPX data.

---

```{r overall-summary}
if (is.null(log) || length(log) == 0) {
  cat("No exclusion steps were recorded in this session.")
  knitr::knit_exit()
}
```

```{r steps-loop, results='asis'}
# Definition of step descriptions for better clarity
step_rationales <- list(
  "A. Data Preview – Exclude samples with QC Warning" = "Measurements flagged by Olink's NPX Manager or NPX Explore with a QC Warning (anything other than 'PASS') are removed at the protein level to prevent unreliable data points from biasing the analysis.",
  "B. Preprocessing > 3. LOD Detection – Filter proteins >50% below LOD" = "Assays where a significant portion of samples (typically >50%) fall below the Limit of Detection (LOD) are excluded. High proportions of missing or below-LOD data can reduce statistical power and lead to false discoveries.",
  "B. Preprocessing > 4. Outlier Detection – Exclude outlier samples" = "Samples identified as multivariate outliers (using UMAP and distance-based heuristics) are excluded. These samples may represent technical failures or biological extremes that do not represent the common sample distribution.",
  "B. Preprocessing > 5. Manual Exclusion" = "User-defined exclusion of specific samples (e.g., control samples) or proteins based on study protocol or prior quality assessment."
)

for (i in seq_along(log)) {
  entry <- log[[i]]
  rationale <- step_rationales[[entry$step]]
  
  # Main Section for each step
  cat("# Step ", i, ": ", entry$step, "\n\n", sep="")
  cat("**Performed on:** ", entry$timestamp, "\n\n")

  if (!is.null(rationale)) {
    cat("## Background\n\n")
    cat(rationale, "\n\n")
  }

  if (!is.null(entry$description) && nchar(entry$description) > 0) {
    cat("## Execution Details\n\n")
    cat(entry$description, "\n\n")
  }

  # Samples section
  if (!is.null(entry$samples_excluded) && length(entry$samples_excluded) > 0) {
    cat("## Samples Excluded (", length(entry$samples_excluded), ")\n\n", sep = "")
    cat(paste(entry$samples_excluded, collapse = ", "), "\n\n")
  }

  # Proteins / assays section
  if (!is.null(entry$proteins_excluded) && length(entry$proteins_excluded) > 0) {
    cat("## Proteins Excluded (", length(entry$proteins_excluded), ")\n\n", sep = "")
    cat(paste(entry$proteins_excluded, collapse = ", "), "\n\n")
  }

  # Dataset before/after
  if (!is.null(entry$before) && !is.null(entry$after)) {
    cat("## Pre/Post Filtering Summary\n\n")
    cat("| Metric | Before | After | Removed |\n")
    cat("|--------|--------|-------|---------|\n")
    if (!is.null(entry$before$samples)) {
      cat("| Samples |", entry$before$samples, "|",
          entry$after$samples, "|",
          entry$before$samples - entry$after$samples, "|\n")
    }
    if (!is.null(entry$before$proteins)) {
      cat("| Proteins |", entry$before$proteins, "|",
          entry$after$proteins, "|",
          entry$before$proteins - entry$after$proteins, "|\n")
    }
    if (!is.null(entry$before$rows)) {
      cat("| Measurements |", entry$before$rows, "|",
          entry$after$rows, "|",
          entry$before$rows - entry$after$rows, "|\n")
    }
    cat("\n")
  }

  # Extra notes
  if (!is.null(entry$notes) && nchar(entry$notes) > 0) {
    cat("## Additional Notes\n\n")
    cat(entry$notes, "\n\n")
  }

  # cat("\\newpage\n\n")
}
```

---

\vspace{20cm}
\textbf{\textcolor{blue}{Report generated by olinkWrappeR on `r Sys.time()`}}

